---
title: "Home Work 5"
author: "Gideon Donkor Bonsu"
date: "07/11/2020"
output: word_document
---

```{r setup, message = FALSE}
library(descr)
load("glsshouseholdSec1_Sec5merged.RData")
```


### QUESTION 1

```{r}

# Probability that there will be no 100-year floods in the 21st century
prob_no100yearFlood <- dpois(0,1); prob_no100yearFlood

# Probability that there will be exactly one 100-year flood in the 21st century
prob_one100yearFlood <- dpois(1,1); prob_one100yearFlood

# Probability that there will be exactly two 100-year flood in the 21st century
prob_two100yearFlood <- dpois(2,1); prob_two100yearFlood

# Probability that there will be three or more 100-year floods in any 100-year period
prob_greaterThan3 <- 1 - ppois(2,1); prob_greaterThan3

```

a) The probability that no 100-year floods will be recorded in the 21st century is 0.3679.

b) The probability that exactly one 100-year flood will be recorded in the 21st century is also 0.3679. However, there is a 0.1839 chance that two 100-year floods will occur.

c) The probability that there will be three or more 100-year floods in any 100-year period is 0.0803.

```{r}
# Create a spike plot of the probability distribution for the number of 100-year floods in a given 100-year.
numberOfOccurences <- c(0:10)
probabilities <- dpois(numberOfOccurences,1)

plot(numberOfOccurences, probabilities,
     main = "Probability Distribution of 100-Year \n Floods In Ghana",
     ylab = "Probability", xlab = "Number of Floods",
     ylim = c(0,0.4))

points(numberOfOccurences, probabilities, pch = 16, col = "red")


# Calculate the mean and standard deviation of the distribution
mean <- sum(numberOfOccurences * probabilities); mean
deviation <-sqrt(sum((numberOfOccurences - mean)^2 * probabilities)); deviation
```
d) The probability distribution for the occurrence of 100-year floods in Ghana is heavily skewed to the right. This is understandable because as the number of occurrences increase over the interval, we expect a lower likelihood. For instance, intuitively, recording ten 100-year floods is very unlikely as compared to observing one occurrence. In the long run, we expect to record one 100-year flood (mean of the distribution) where the number of possible floods deviates from this value by 1 (standard deviation).


### QUESTION 2
```{r}
# The probability that the next call will arrive within 5 minutes (note: 5 minutes = 5/60 hours) ð‘ƒ(ð‘‹ â‰¤ 5/60). 
p_within5Mins <- pexp(5/60, 10); p_within5Mins

# The probability that the next call will arrive within 10 minutes
p_within10mins <- pexp(10/60, 10); p_within10mins

# Given a mean rate of 25 calls per hour,the probability that the next call will arrive within 5 minutes.
p_within5Mins1 <- pexp(5/60, 25); p_within5Mins1

# Given a mean rate of 25 calls per hour,the probability that the next call will arrive within 10 minutes.
p_within10mins1 <- pexp(10/60, 25); p_within10mins1


# Two values between which there is a 95% probability of arrival times between calls.
lowerBound <- qexp(0.025, 25); lowerBound
upperBound <- qexp(0.975,25); upperBound


```
a) At a mean rate of 10 calls per hour, the probability that the next call will arrive within 5 minutes is 0.5654.

b) At a mean rate of 10 calls per hour, the probability that the next call will arrive within 5 minutes is 0.8111.

c) Given a new mean rate of 25 calls per hour,the likelihood that the next call will arrive within 5 minutes is 0.5654. On the other hand, there is a 0.9845 chance that the next call will arrive within 10 minutes.

d) With calls now arriving at a mean rate of 25 per hour, 95% probability arrival times between calls will lie between 0.0010 hours and 0.1476 hours.




### QUESTION 3
```{r}
# Confidence interval for the population mean
sample_mean <-15.3 
s <- 3.2
n <- 125;

# Since the sample size is 125, using z* provides a good approximation
margin_of_error <- 1.96 * (s/sqrt(n)); margin_of_error
lowerBound1 <- sample_mean - margin_of_error; lowerBound1
upperBound1 <- sample_mean + margin_of_error; upperBound1


# Confidence interval for the population proportion.
p <- 83/125
margin_of_error1 <- 1.96 * (sqrt((p *(1-p))/n)); margin_of_error1
lowerBound2 <- p - margin_of_error1; lowerBound2
upperBound2 <- p + margin_of_error1; upperBound2


```

a) The interval is between 14.74 and 15.86.

b) We are 95% confident that the average weekly television viewing time for all residents of Greater Accra aged 55 or older is between 14.74 and 15.86.

c) The interval is between 0.58 and 0.75.

d) We are 95% confident the proportion of Greater Accra residents age 55 years and over who watch the evening news at least three weeknights per week is between 58.12% and 74.68%.


### QUESTION 4
```{r}
# Step 2: Collect and summarize the data
# Random sample greater than 30.
x <- c(5.65, 5.44, 5.42, 5.42, 5.53, 5.34, 5.54, 5.45, 5.52, 5.41, 5.57, 5.40, 5.53, 5.54, 5.55, 5.62, 5.56, 5.46, 5.44, 5.51, 5.47, 5.40, 5.47, 5.61, 5.53, 5.32, 5.67, 5.29, 5.49, 5.55, 5.77, 5.57, 5.42, 5.58, 5.58, 5.50, 5.32, 5.50, 5.53, 5.58, 5.61, 5.45, 5.44, 5.25, 5.56, 5.63, 5.50, 5.57, 5.67, 5.36)

# Checking if the distribution is normal.
hist(x, main = "Weight of Tea Per Bag in a Factory", xlab = "Amount (in grams)")

# Calculating the sample mean and sample standard deviation
x_bar <- mean(x); x_bar
s <- sd(x); s

# Step 3: Assess the evidence
t.test(x, mu = 5.5, alternative = "two.sided")
```
i) **Step 1: State the claim**: 
  Null Hypothesis: The mean weight of tea in a bag filled by the machine is 5.5 grams(i.e. Ho: mu = 5.5grams).

  Alternate Hypothesis: The mean weight of tea in a bag filled by the machine is different from 5.5 grams.(i.e. Ha: mu != 5.5grams).

ii) **Step 2: Collect and summarize the data**: 
 The sample mean is 5.5018 and its sample standard deviation is 0.1054764. 
 We will use sample standard deviation to estimate the population standard deviation, which introduces greater uncertainty, therefore we have to conduct a t-test for the population mean.
Again, we will have to assume that the bags of tea were randomly selected.
  Finally, the sample is over 30 and the distribution, although doesn't matter in this case, is fairly normal as seen in the histogram. 
All the required conditions are met, hence, we can carry on with the test.

iii) **Step 3: Assess the evidence**:     
From the above calculations, the test statistic is 0.12067 with an associated p-value of  0.9044.

iv) **Make conclusion**:
  At the 0.05 level of significance, the p-value is large (0.9044 > 0.05) which means the data is not significant. Therefore, the data does not provide enough evidence to reject the claim that the mean weight of tea in a bag filled by the machine is 5.5 grams. Hence, we cannot conclude that the average weight of tea per bag is different from 5.5 grams.
  
  We are susceptible to making a type 2 error. This is because we failed to reject the null hypothesis which can turn out to be true.



### QUESTION 5
```{r}
mean <- 1068
s <-  204
# The probability that a select student's score on the SAT is at least 1130
p_atLeast1330 <- 1 - pnorm(1130, mean, s); p_atLeast1330

# The probability that the mean score of 25 selected graduates is at least 1130
p_atLeast1330s <- 1 - pnorm(1130, mean, s/sqrt(25)); p_atLeast1330s

```
a) If 1 high school graduate is randomly selected, the probability that his or her score on the SAT is at least 1130 is 0.3806.

b) If 25 high school graduates are randomly selected, find the probability that their mean score is at least 1130 is 0.0643.

**Step 1: State the claim**:
Null Hypothesis: High school graduates who received the Columbia Review Course will have a mean score of 1068 (i.e. Ho: mu = 1068).

Alternate Hypothesis: High school graduates who received the Columbia Review Course will have a mean score higher than 1068 (mu > 1068).

**Step 2: Collect and summarize the data**:
The population standard deviation for test scores for high school graduates who took the SAT in 2018 during their final year of high school is known. Hence, we can conduct a z-test. 
Again, the distribution of the test scores is known to be normal and the high school graduates used for the study were randomly selected.
The required conditions are therefore met, hence, we can proceed with the test.

**Step 3: Assess the evidence**:
```{r}
# Calculating the test statistic
s <- 204/sqrt(25)
diff_m <- 1130 - 1068
z_test <- (diff_m/s);z_test

# P-Value
pnorm(z_test, lower.tail = FALSE)

```
From the above calculations, the test statistic is 1.5196 with an associated p-value of  0.0643.

**Make conclusion**:
  At 0.10 level of significance, the p-value is small (0.0643 < 0.10) which means the data is significant. Therefore, it does provide enough evidence to reject the claim that high school graduates who received the Columbia Review Course will have a mean score of 1068. Hence, we can conclude that indeed the average SAT score of graduates who received the Columbia Review Course is higher than the national average.

d)  We are susceptible to making a type 1 error. This is because we rejected the null hypothesis which can turn out to be true.

### QUESTION 6
```{r}


# Creating a variable subset with the variables s2aq3 and loc2
variable_subset <- subset(glsshouseholdSec1_Sec5merged, select=c("s2aq3","loc2"))


# Creating a new variable that is equal to 0 if the respondent did not attain a secondary school qualification and is equal to 1 if the respondent did attain a secondary school qualification or higher. 
variable_subset $ education <- rep(NA, nrow(variable_subset))
variable_subset $ education[variable_subset $ s2aq3 < 3] <- 0
variable_subset $ education[variable_subset $ s2aq3 >= 3] <- 1

# Collecting and summarize the data for Hypothesis Testing
freq(variable_subset$education)
p_hat <- 0.153
p <- 0.147
sample_size <- 42140

# Checking if conditions for using a z-test are met
sample_size * p >= 10
sample_size *(p *(1-p)) >= 10

# Finding the test statistic and p-value
zTest <- (p_hat - p)/sqrt((p * (1-p))/sample_size); zTest
pValue <- pnorm(zTest, lower.tail = FALSE) + pnorm(-zTest); pValue


# Constructing a 95% confidence interval estimate for the proportion of Ghanaians age 15 and older who had a secondary school qualification or higher in 2018.
lower <- p_hat - (2 *  sqrt((p_hat * (1-p_hat))/sample_size)); lower
upper <- p_hat + (2 *  sqrt((p_hat * (1-p_hat))/sample_size)); upper
```
**Step 1: State the claim**: 
Null Hypothesis : The percentage of Ghanaians age 15 and older with a secondary school qualification or higher has not changed since 2013. (Ho: Po = 0.147)

Alternate Hypothesis : The percentage of Ghanaians age 15 and older with a secondary school qualification or higher has changed since 2013. (Ho: Po != 0.147)

**Step 2: Collect and summarize the data**:
Sample proportion is 0.153. Since (42140 * 0.147) >= 10 and (42140 * (1-0.147)) >= 10, we can proceed with the z-test.

**Step 3: Assess the evidence**: 
While the test statistic is 46.13072, its p-value is 0.0005

 **Make conclusion**:
  At the 0.05 level of significance, the p-value is very small (0.0005 < 0.05) which means the data is significant. Therefore, it does provide enough evidence to reject the claim that the percentage of Ghanaians age 15 and older with a secondary school qualification or higher has not changed since 2013. Hence, we can conclude that there has been changes to that percentage since 2013.

d) Since we failed to reject the null hypothesis, we are susceptible to making a type 2 error in this case. 

e) At 95% confidence level, the interval is between 0.1495 and 0.1565


### QUESTION 7
a)
```{r}

# Generate a table for the proportion of Ghanaians, age 15 and over, who have attained a secondary school qualification or higher.(explanatory var = columns)
t1 <- table(variable_subset $ education, variable_subset $ loc2); t1

# convert to a table of proportions, 2 specifies column proportions
t1_prop <- prop.table(t1,2); t1_prop

# Generate a barplot for the proportion of Ghanaians, age 15 and over, who have attained a secondary school qualification or higher."
barplot(t1_prop[2, ], 
        main = " Proportion of Young Ghanaians Who Have \n a Secondary School Qualification or Higher.",
        names = c("Urban\nYouth", "Rural\nYouth"),
        ylab = "Have an SHS Qualification or Higher.",
        ylim=c(0,0.30), col = "#3c1361")


```

Given that 35% urban Ghanaians, age 15 and over, attained a secondary school qualification or higher as opposed to approximately 16% of rural Ghanaians, the graph suggest that there is a possible association between urban vs rural location and educational attainment in Ghana.















